---
output: pdf_document
---

```{r setup, include=FALSE, message = FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(tidyr)
library(stringr)
library(egg)
```

```{r bias visual, echo = FALSE, message = FALSE, warning = FALSE}
bias_df <- read.csv('drcrbias.csv') %>%
  filter(`Intervention` == 'Levadopa') %>%
  select(-`Intervention`)
  
df_dropped <- bias_df %>% select(-Type)
df_melted <- df_dropped %>%
  gather(key = "Bias_Type", value = "Bias_Level", -Study)

overall_bolder = c(rep(0, 7), 1)

color_map <- c('Low' = '#00D811', 'Moderate' = '#FFDC36', 'Serious' = 'red', 'Critical' = 'darkred')

bias <- ggplot(df_melted, aes(x = Bias_Type, y = Study, fill = Bias_Level)) +
  geom_tile(color = "white", size = 0.5) +
  scale_fill_manual(values = color_map, limits = c('Low', 'Moderate', 'Serious', 'Critical')) +
  geom_vline(xintercept = 7.501, linetype = 'solid') + 
  geom_vline(xintercept = 7.499, linetype = 'solid') +
  labs(title = 'ROBINS-I Tool Analysis: Levadopa', x = 'Bias Domain', y = NULL) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(2)),  
    axis.title.x = element_text(size = rel(1.3)),
    axis.title.y = element_text(size = rel(1.3)),
    axis.text.x = element_text(face = ifelse(overall_bolder, 'bold', 'plain')),
    legend.text = element_text(size = rel(1.1))
    )

bias
