---
title: '\textbf{Metformin and AMD: Review of Current Literature}'
output:
  pdf_document: default
---

```{r setup, echo=FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(ggrepel)
library(tidyr)
library(stringr)
library(egg)
```

\definecolor{linkblue}{rgb}{0,0,0.7}

\begin{center}
    \textbf{\Large Prevalence, Clinical Trial Count, and DRCR Protocol Count}
\end{center}

```{r Dictionary Generation, echo = FALSE}
prevalence_dict <- c(DR = 0.285 * 0.148, DME = 0.027 * 0.148 * 0.148, ERM = 0.026, VMT = 0.0107, RR = NULL, MH = 0.0002, Wet_AMD = 0.0094, Dry_AMD = 0.1164)

clinical_dict <- c(DR = 614, DME = 578, ERM = 86, VMT = 18, RR = 11, MH = 113, Wet_AMD = 512, Dry_AMD = 67)
```

```{r Protocols, Prevalence, Clinical Trials, fig.width=4, fig.height=7, warning=FALSE, echo = FALSE}
protocol_df <- read.csv('drcrprotocols.csv')
protocol_df <- subset(protocol_df, Disease != 'MISC')

formatted_prevalence_df <- protocol_df %>%
  group_by(Disease) %>%
  summarise(Number_of_Rows = n())

dry_amd_row = data.frame(Disease = 'Dry_AMD', Number_of_Rows = 0)
formatted_prevalence_df <- rbind(formatted_prevalence_df, dry_amd_row)
formatted_prevalence_df$Prevalence <- prevalence_dict[match(formatted_prevalence_df$Disease, names(prevalence_dict))]
formatted_prevalence_df$Clinical_Trials_Count <- clinical_dict[match(formatted_prevalence_df$Disease, names(clinical_dict))]

formatted_prevalence_df <- formatted_prevalence_df %>%
  rename(DRCR_Protocol_Count = Number_of_Rows)
formatted_prevalence_df <- formatted_prevalence_df %>%
  mutate(Disease = gsub("_", " ", Disease))

common_theme <- theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(1.1), hjust = 0.5),  
    axis.title.x = element_text(size = rel(0.8)),
    axis.title.y = element_text(size = rel(1))
  )

plt1 <- ggplot(formatted_prevalence_df, aes(x = Prevalence, y = DRCR_Protocol_Count, label = Disease)) +
  geom_point(size = 1.5) +
  geom_text_repel(box.padding = 0.3, segment.color = NA, size = 2.5, family = "serif") +
  labs(title = 'Prevalence vs DRCR Protocol Count',
       x = 'Prevalence',
       y = 'DRCR Protocol Count') + 
  common_theme 

plt2 <- ggplot(formatted_prevalence_df, aes(x = Prevalence, y = Clinical_Trials_Count, label = Disease)) + 
  geom_point(size = 1.5) +
  geom_text_repel(box.padding = 0.3, segment.color = NA, size = 2.5, family = "serif") +  
  labs(title = NULL,  # Remove the title from labs
       x = 'Prevalence', 
       y = 'Clinical Trial Count') + 
  ggtitle(str_wrap('Prevalence vs Clinical Trial Count (as listed on clinicaltrials.gov)', width = 60)) +
  common_theme

ggarrange(plt1, plt2, nrow = 2)

```

\textit{\small Note: Prevalence values are representative of adults ages 40 or older}
\
\textit{\small Note: Insufficient data prevents a conclusive assessment of radiation retinopathy prevalence}

\begin{minipage}[t]{0.65\textwidth}
  \subsection*{\small Prevalence Values Literature Review:}
  \begin{itemize}
    \small
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9634594/}{\textcolor{linkblue}{Age Related Macular Degeneration (AMD) (Wet and Dry)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2945293/}{\textcolor{linkblue}{Diabetic Retinopathy (DR)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2945293/}{\textcolor{linkblue}{Diabetic Macular Edema (DME)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5623383/}{\textcolor{linkblue}{Epiretinal Membrane (ERM)}}
      \item \href{https://pubmed.ncbi.nlm.nih.gov/28806215/}{\textcolor{linkblue}{Vitreomacular Traction (VMT)}}
      \item \href{https://jamanetwork.com/journals/jamaophthalmology/fullarticle/2603482}{\textcolor{linkblue}{Macular Hole (MH)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10093581/}{\textcolor{linkblue}{Radiation Retinopathy (RR)}}
  \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
  \subsection*{\small Number of Clinical Trials:}
  \begin{itemize}
    \small
     \item \href{https://www.clinicaltrials.gov}{\textcolor{linkblue}{Clinical Trials Website}}
     \item \href{https://github.com/Matthewdhuh/drcrMetaReview}{\textcolor{linkblue}{Data and Code Repository}}
  \end{itemize}
\end{minipage}

\section*{\Large \centering Metformin's Effect on AMD: ROBINS-I Tool Bias Analysis}


```{r Metformin Effect on AMD Forest Plot, warning = FALSE, echo = FALSE, message = FALSE}

drcrrates <-read.csv('drcrrates.csv')
study_order <- c('Jiang (2022)', 'Chen (2019)', 'Tseng (2023)', 'Gokhale (2023)', 'Eton (2022)',
                 'Stewart (2020)',
                 'Starr (2022)', 'Brown (2019)', 'Aggarwal (2024)', 'Blitzer (2021)', 'Khanna (2022)', 'Kaufmann (2023)', 'Lee (2019)')

type_order <- c('Case control', 'Cross sectional', 'Retrospective cohort')


drcrrates$Paper <- factor(drcrrates$Paper, levels = study_order, ordered = TRUE)
drcrrates$Type <- factor(drcrrates$Type, levels = type_order, ordered = TRUE)

common_theme <- theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(2)),  
    axis.title.x = element_text(size = rel(1.3)),
    axis.title.y = element_text(size = rel(1.3))
  )

rates <- ggplot(drcrrates, aes(x = Rate.ratio, y = Paper, color = Type)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper, height = 0.2), position = position_dodge(width = 0.5)) +
  labs(title = 'Metformin Rate Ratios vs AMD',
       x = 'Rate Ratio',
       y = NULL,
       color = 'Study Type') +
  geom_vline(xintercept = 1, linetype = 'dashed', color = 'red') +
  common_theme +
  theme(legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = rel(1.1)),
        ) +
  scale_color_manual(values = c('blue','purple', '#1EF0E2'))

```

```{r ROBINS I Tool Bias Plot, warning = FALSE, echo = FALSE}

bias_df <- read.csv('drcrbias.csv') %>%
  filter(`Intervention` == 'Metformin') %>%
  select(-`Intervention`)
  
bias_df$Study <- factor(bias_df$Study, levels = study_order, ordered = TRUE)
df_dropped <- bias_df %>% select(-Type)
df_melted <- df_dropped %>%
  gather(key = "Bias_Type", value = "Bias_Level", -Study)

overall_bolder = c(rep(0, 7), 1)

color_map <- c('Low' = '#00D811', 'Moderate' = '#FFDC36', 'Serious' = 'red', 'Critical' = 'darkred')

bias <- ggplot(df_melted, aes(x = Bias_Type, y = Study, fill = Bias_Level)) +
  geom_tile(color = "white", size = 0.5) +
  scale_fill_manual(values = color_map, limits = c('Low', 'Moderate', 'Serious', 'Critical')) +
  geom_vline(xintercept = 7.501, linetype = 'solid') + 
  geom_vline(xintercept = 7.499, linetype = 'solid') +
  labs(title = 'ROBINS-I Tool Analysis', x = 'Bias Domain', y = NULL) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(2)),  
    axis.title.x = element_text(size = rel(1.3)),
    axis.title.y = element_text(size = rel(1.3)),
    axis.text.x = element_text(face = ifelse(overall_bolder, 'bold', 'plain')),
    legend.text = element_text(size = rel(1.1))
    )
```

```{r Arrange Bias and Rates, fig.width = 10, fig.height = 5, warning = FALSE, echo = FALSE}
bias <- bias +
  labs(title = NULL, x = 'Bias Domain', y = NULL) + scale_y_discrete(position = "right")

rates <- rates + labs(title = NULL, x = 'Rate Ratio', y = NULL)
plot <- ggarrange(rates, bias, ncol = 2)
```




\begin{minipage}[t]{0.5\textwidth}
    \begin{flushleft}
      \subsection*{\small Bias Domains:}
        \small
        {\textbf{D1:} Bias due to confounding} \\
        {\textbf{D2:} Bias due to selection of participants} \\
        {\textbf{D3:} Bias in classification of interventions} \\
        {\textbf{D4:} Bias due to deviation from intended interventions} \\
        {\textbf{D5:} Bias due to missing data} \\
        {\textbf{D6:} Bias in measurement of outcomes} \\
        {\textbf{D7:} Bias in selection of the reported result} \\
    \end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
  \begin{flushleft}
    \subsection*{\small Major limitations:}
    \small
      \href{https://pubmed.ncbi.nlm.nih.gov/31250582/}{\textcolor{blue}{\text{Lee:}}} Case-control (selection bias) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/37589984/}{\textcolor{blue}{\text{Kaufmann:}}} Case-control (selection bias) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/38259182/}{\textcolor{blue}{\text{Khanna:}}} Case-control (selection bias) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/33475696/}{\textcolor{blue}{\text{Blitzer:}}} Case-control (selection bias) \\ 
      \href{https://pubmed.ncbi.nlm.nih.gov/38019527/}{\textcolor{blue}{\text{Aggarwal:}}} Case-control (selection bias) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/30973575/}{\textcolor{blue}{\text{Brown:}}} Case-control (selection bias) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/36151313/}{\textcolor{blue}{\text{Starr:}}} Case-control (selection bias) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/32525055/}{\textcolor{blue}{\text{Stewart:}}} Cross-sectional (no temporal information) \\
      \href{https://pubmed.ncbi.nlm.nih.gov/33607930/}{\textcolor{blue}{\text{Eton:}}} Unclear methods on time-dependent analysis \\
      \href{https://pubmed.ncbi.nlm.nih.gov/35115301/}{\textcolor{blue}{\text{Gokhale:}}} Young population (greater than 40 yr); short time frame for time-dependent analysis \\
      \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9963185/}{\textcolor{blue}{\text{Tseng:}}} No time-dependent analysis \\
      \href{https://pubmed.ncbi.nlm.nih.gov/31781371/}{\textcolor{blue}{\text{Chen:}}} No time-dependent analysis \\
      \href{https://pubmed.ncbi.nlm.nih.gov/35473747/}{\textcolor{blue}{\text{Jiang:}}} No time-dependent analysis \\
  \end{flushleft}
\end{minipage}


\




\small \textbf{ROBINS-I (Risk Of Bias In Non-randomized Studies – of Interventions)} is used by the Cochrane Scientific Committee for assessing bias in non-randomized studies of interventions. [\href{https://methods.cochrane.org/robins-i}{\textcolor{linkblue}{ROBINS-I Tool homepage}}]





