---
title: '\textbf{Metformin and AMD: Review of Current Literature}'
author: '\textit{Matthew Huh, BA}'
output:
  pdf_document: default
---

```{r setup, echo=FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(ggrepel)
library(egg)
library(tidyr)
library(grid)
```

\definecolor{linkblue}{rgb}{0,0,0.7}

\begin{center}
    \textbf{\Large Prevalence, Clinical Trial Count, and DRCR Protocol Count}
\end{center}

```{r Dictionary Generation, echo = FALSE}
prevalence_dict <- c(DR = 0.285 * 0.148, DME = 0.38 * 0.148, ERM = 0.026, VMT = 0.0107, RR = NULL, MH = 0.0002, Wet_AMD = 0.0094, Dry_AMD = 0.1164)

clinical_dict <- c(DR = 614, DME = 578, ERM = 86, VMT = 18, RR = 11, MH = 113, Wet_AMD = 512, Dry_AMD = 67)
```

```{r Protocols, Prevalence, Clinical Trials, fig.width=5/1.2, fig.height=8/1.2, warning=FALSE, echo = FALSE}
protocol_df <- read.csv('drcrprotocols.csv')
protocol_df <- subset(protocol_df, Disease != 'MISC')

formatted_prevalence_df <- protocol_df %>%
  group_by(Disease) %>%
  summarise(Number_of_Rows = n())

dry_amd_row = data.frame(Disease = 'Dry_AMD', Number_of_Rows = 0)
formatted_prevalence_df <- rbind(formatted_prevalence_df, dry_amd_row)
formatted_prevalence_df$Prevalence <- prevalence_dict[match(formatted_prevalence_df$Disease, names(prevalence_dict))]
formatted_prevalence_df$Clinical_Trials_Count <- clinical_dict[match(formatted_prevalence_df$Disease, names(clinical_dict))]

formatted_prevalence_df <- formatted_prevalence_df %>%
  rename(DRCR_Protocol_Count = Number_of_Rows)
formatted_prevalence_df <- formatted_prevalence_df %>%
  mutate(Disease = gsub("_", " ", Disease))

common_theme <- theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(1.1)),  
    axis.title.x = element_text(size = rel(0.8)),
    axis.title.y = element_text(size = rel(1))
  )

plt1 <- ggplot(formatted_prevalence_df, aes(x = Prevalence, y = DRCR_Protocol_Count, label = Disease)) +
  geom_point(size = 1.5) +
  geom_text_repel(box.padding = 0.3, segment.color = NA, size = 2.5, family = "serif") +
  labs(title = 'Prevalence vs DRCR Protocol Count',
       x = 'Prevalence',
       y = 'DRCR Protocol Count') + 
  common_theme 

plt2 <- ggplot(formatted_prevalence_df, aes(x = Prevalence, y = Clinical_Trials_Count, label = Disease)) + 
  geom_point(size = 1.5) +
  geom_text_repel(box.padding = 0.3, segment.color = NA, size = 2.5, family = "serif") +  
  labs(title = 'Prevalence vs Clinical Trial Count', 
       x = 'Prevalence', 
       y = 'Clinical Trial Count') + 
  common_theme 

ggarrange(plt1, plt2)


```

\textit{\small Note: Insufficient data prevents a conclusive assessment of Radiation Retinopathy prevalence}

\begin{minipage}[t]{0.65\textwidth}
  \subsection*{\small Prevalence Values Literature Review:}
  \begin{itemize}
    \small
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC9634594/}{\textcolor{linkblue}{Age Related Macular Degeneration (AMD) (Wet and Dry)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2945293/}{\textcolor{linkblue}{Diabetic Retinopathy (DR)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2945293/}{\textcolor{linkblue}{Diabetic Macular Edema (DME)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5623383/}{\textcolor{linkblue}{Epiretinal Membrane (ERM)}}
      \item \href{https://pubmed.ncbi.nlm.nih.gov/28806215/}{\textcolor{linkblue}{Vitreomacular Traction (VMT)}}
      \item \href{https://jamanetwork.com/journals/jamaophthalmology/fullarticle/2603482}{\textcolor{linkblue}{Macular Hole (MH)}}
      \item \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10093581/}{\textcolor{linkblue}{Radiation Retinopathy (RR)}}
  \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
  \subsection*{\small Number of Clinical Trials:}
  \begin{itemize}
    \small
     \item \href{https://www.clinicaltrials.gov}{\textcolor{linkblue}{Clinical Trials Website}}
     \item \href{https://github.com/Matthewdhuh/drcrMetaReview}{\textcolor{linkblue}{Data and Code Repository}}
  \end{itemize}
\end{minipage}

\section*{\Large \centering Metformin's Effect on AMD: ROBINS-I Tool Bias Analysis}


```{r Metformin Effect on AMD Forest Plot, warning = FALSE, echo = FALSE, message = FALSE}

drcrrates <-read.csv('drcrrates.csv')
study_order <- c('Brown (2019)', 'Aggarwal (2024)', 'Blitzer (2021)', 'Kaufmann (2023)', 'Lee (2019)',
                 'Jiang (2022)', 'Chen (2019)', 'Starr (2022)', 'Gokhale (2023)', 'Eton (2022)',
                 'Stewart (2020)', 'Khanna (2022)')

type_order <- c('Cross sectional', 'Retrospective cohort', 'Case control')


drcrrates$Paper <- factor(drcrrates$Paper, levels = study_order, ordered = TRUE)
drcrrates$Type <- factor(drcrrates$Type, levels = type_order, ordered = TRUE)

common_theme <- theme_minimal() +
  theme(
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(2)),  
    axis.title.x = element_text(size = rel(1.3)),
    axis.title.y = element_text(size = rel(1.3))
  )

rates <- ggplot(drcrrates, aes(x = Rate.ratio, y = Paper, color = Type)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper, height = 0.2), position = position_dodge(width = 0.5)) +
  labs(title = "Metformin's Effect on AMD",
       x = 'Rate Ratio',
       y = NULL,
       color = 'Study Type') +
  geom_vline(xintercept = 1, linetype = 'dashed', color = 'red') +
  common_theme +
  theme(legend.position = 'bottom',
        legend.title = element_blank(),
        legend.text = element_text(size = rel(1.1))) +
  scale_color_manual(values = c('#FFDC36', '#00D811', 'red'))

```

```{r ROBINS I Tool Bias Plot, warning = FALSE, echo = FALSE}

bias_df = read.csv('drcrbias.csv')

bias_df$Study <- factor(bias_df$Study, levels = study_order, ordered = TRUE)
df_dropped <- bias_df %>% select(-Type)
df_melted <- df_dropped %>%
  gather(key = "Bias_Type", value = "Bias_Level", -Study)

color_map <- c('Low' = '#00D811', 'Moderate' = '#FFDC36', 'Serious' = 'red', 'Critical' = 'darkred')

bias <- ggplot(df_melted, aes(x = Bias_Type, y = Study, fill = Bias_Level)) +
  geom_tile(color = "white", size = 0.5) +
  scale_fill_manual(values = color_map, limits = c('Low', 'Moderate', 'Serious', 'Critical')) +
  geom_vline(xintercept = 7.501, linetype = 'solid') + 
  geom_vline(xintercept = 7.499, linetype = 'solid') +
  labs(title = 'ROBINS-I Tool Bias Analysis', x = 'Bias Domain', y = NULL) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.x = element_text(face = "plain"),
    text = element_text(family = "serif"),
    plot.title = element_text(size = rel(2)),  
    axis.title.x = element_text(size = rel(1.3)),
    axis.title.y = element_text(size = rel(1.3)),
    legend.text = element_text(size = rel(1.1)) 
  )
```

```{r Arrange Bias and Rates, fig.width = 10, fig.height = 5, warning = FALSE, echo = FALSE}
bias <- bias +
  labs(title = 'ROBINS-I Tool Analysis', x = 'Bias Domain', y = NULL) + 
  theme(axis.text.y = element_blank())
rates <- rates + labs(title = 'Metformin Rate Ratios vs AMD', x = 'Rate Ratio', y = NULL)
plot <- ggarrange(rates, bias, ncol = 2)
```

\


\begin{minipage}[t]{0.57\textwidth}
    \subsection*{\small ROBINS-I Tool Homepage:}
    \begin{itemize}
      \small
        \item \href{https://methods.cochrane.org/robins-i}{\textcolor{linkblue}{ROBINS-I Tool Homepage}}
    \end{itemize}
\end{minipage}
\hfill
\begin{minipage}[t]{0.55\textwidth}
    \begin{flushleft}
        \small
        \textit{D1: Bias due to confounding} \\
        \textit{D2: Bias due to selection of participants} \\
        \textit{D3: Bias in classification of interventions} \\
        \textit{D4: Bias due to deviation from intended interventions} \\
        \textit{D5: Bias due to missing data} \\
        \textit{D6: Bias in measurement of outcomes} \\
        \textit{D7: Bias in selection of the reported result} \\
        \textit{Overall: Overall bias of study}
    \end{flushleft}
\end{minipage}






